### 2017-12-14 Paper/cover.tex
added outline for the Paper

%-------------------COVER-------------------------------------------
\begin{center}
\begin{Large}
{\bf Bayesian Guided Optimal Experimental Design for Reactive Multilayers}\\
\end{Large}

\bigskip
%\bigskip

Manav Vohra,$^1$ Xun Huan,$^2$ Omar M. Knio$^{1,3}$ \\

\bigskip
\bigskip

\normalsize
$^1$Department of Mechanical Engineering and Materials Science\\
Duke University\\
Durham, NC 27708\\

\bigskip

$^2$Department of Aeronautics and Astronautics\\
Massachusetts Institute of Technology\\
Cambridge, MA 02139\\

\bigskip

$^3$Computer, Electrical and Mathematical Sciences and Engineering Division \\
King Abdullah University of Science and Technology \\
Thuwal 23955-6900, Kingdom of Saudi Arabia\\

\end{center}

\vspace{6cm}

\begin{tabbing}
Corresponding Author: \hspace{5mm} \= Omar M. Knio\\
       \>  Department of Mechanical Engineering and Materials Science\\
       \>  Duke University\\
       \>  144 Hudson Hall, Box 90300 \\
       \>  Durham, NC 27708 \\
       \> \\
Phone: \> (919) 660-5344 \\
Fax:   \> (919) 660-8963 \\
Email: \>  omar.knio@duke.edu   \\
\\
Submitted to: \> \textit{Journal} \\
\> December 2015\\

\end{tabbing}
\clearpage