### 2017-12-19 Paper/cover.tex
moved cover to a separate file

\begin{center}
\textsc{
Derivative-based global sensitivity approach to 
efficient surrogate modeling with application to molecular dynamics
}

\bigskip

\bigskip

Manav Vohra, Alen Alexanderian, Sankaran Mahadevan

\end{center}

\vspace{10cm}

\begin{tabular}{ll}
Corresponding Author:  & TBA \\
                       & Vanderbilt University \\ 
                       & Nashville, TN\\ 
                       & Phone: \\ 
                       & Email: 
\end{tabular}

\bigskip

Submitted to: \emph{Journal of Computational Physics}

\clearpage