### 2017-12-19 Paper/DGSM_PCE_MD_paper.tex
moved intro to a separate file

 \documentclass[letter,12pt]{article}
%
\input{def}
\usepackage[table]{xcolor}
\usepackage{relsize}
\usepackage{bm}
\usepackage{mathrsfs}

 \def\ol{\overline}
 \def\no{\noindent}
 \def\qd{\dot{Q}}

%%%%%%%FLOW DIAGRAM%%%%%%%%%%
\usepackage[latin1]{inputenc}
\usepackage{tikz}
%\usepackage[table]{xcolor}
\usetikzlibrary{shapes,arrows}
% Define block styles
\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
   text width=5.0em, text badly centered, node distance=3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, fill=cyan!20, 
   text width=9.0em, text centered, rounded corners]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=3cm,
   minimum height=2em]
%%%%%%%%%%%%%%%%%%%%%%%%%%

%-------------------------------------------------------------------
\begin{document}

\input{cover}

\tableofcontents

\input{intro}

\section{Background}
\subsection{DGSM}
\[ 
    \mu_i = E\left(\left( \frac{\partial \mathcal{G}}{\partial \theta_i}\right)^2\right)
          \approx \hat\mu_i := \frac1N\sum_{j = 1}^N 
                  \left(\frac{\partial \mathcal{G}(x_j)}{\partial \theta_i}\right)^2
\]
Can consider variance of $\hat\mu_i$.

\subsection{Polynomial chaos}
\section{Method}

\section{Motivating examples}
\subsection{Borehole: Screening, Convergence, Verification}
\subsection{Semilinear elliptic PDE: Screening, Convergence, Verification}
\subsection{Nonlinear Oscillator: Screening, Convergence, Verification}

\section{Application: Phonon Transport in a Si bar}

Seven uncertain parameters; well-established nominal values exist;
\[
   \theta_i \sim U(\bar\theta_i - \epsilon_i, \bar\theta_i + \epsilon_i), 
   \quad \epsilon_i = 0.1 \bar\theta_i. 
\]
\begin{enumerate}
\item Include all details pertaining to the MD simulation: system, simulation, 
ensembles, etc. 
\item Results
\end{enumerate}

\section{Discussion}
\begin{enumerate}
\item Methodology is agnostic to the choice of surrogate.
\item As required for a PCE, the uncertain parameters need not be independent. 
\item Strategy depends on the application - methodology could be implemented 
sequentially or adaptively. 
\end{enumerate}

\end{document}