### 2017-12-19 Paper/DGSM_PCE_MD_paper.tex
separate space for abstract

 \documentclass[letter,12pt]{article}
%
\input{def}
\usepackage[table]{xcolor}
\usepackage{relsize}
\usepackage{bm}
\usepackage{mathrsfs}

 \def\ol{\overline}
 \def\no{\noindent}
 \def\qd{\dot{Q}}

%%%%%%%FLOW DIAGRAM%%%%%%%%%%
\usepackage[latin1]{inputenc}
\usepackage{tikz}
%\usepackage[table]{xcolor}
\usetikzlibrary{shapes,arrows}
% Define block styles
\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
   text width=5.0em, text badly centered, node distance=3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, fill=cyan!20, 
   text width=9.0em, text centered, rounded corners]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=3cm,
   minimum height=2em]
%%%%%%%%%%%%%%%%%%%%%%%%%%

%-------------------------------------------------------------------
\begin{document}

\thispagestyle{empty}
\input{cover}

\baselineskip=22pt

\tableofcontents

\input{abstract}
\clearpage

\input{intro}
\bigskip
\bigskip

\input{bg}
\bigskip
\bigskip

\input{method}
\bigskip
\bigskip

\input{examples}
\bigskip
\bigskip

\input{app}
\bigskip
\bigskip

\input{disc}
\bigskip
\bigskip

\input{ack}
\bigskip
\bigskip

\bibliographystyle{unsrt}
\bibliography{REFER}

\clearpage

\input{figures}

\end{document}