### 2018-04-28 Paper/cover.tex
files added

\begin{center}
\textsc{
Derivative-based global sensitivity approach to efficient surrogate modelling
with Polynomial Chaos
}

\bigskip 
\bigskip

Manav Vohra$^{1}$, Alen Alexanderian$^{2}$, Cosmin Safta$^{3}$, Sankaran Mahadevan$^{1}$

\bigskip
\bigskip

\normalsize
$^1$Department of Civil and Environmental Engineering\\
Vanderbilt University\\
Nashville, TN 37235\\

\bigskip

$^2$Department of Mathematics\\
North Carolina State University\\
Raleigh, NC 27695\\

\bigskip

$^3$Sandia National Laboratories\\
Livermore, CA 94550\\

\end{center}

\vspace{6cm}

\begin{tabbing}
Corresponding Author: \hspace{5mm} \= Sankaran Mahadevan\\
       \>  Department of Civil and Environmental Engineering\\
       \>  Vanderbilt University\\
       \>  272 Jacobs Hall, VU Mailbox: PMB 351831 \\
       \>  Nashville, TN 37235 \\
       \> \\
Phone: \> (615) 322-3040 \\
Fax:   \> (615) 343-3773 \\
Email: \>  sankaran.mahadevan@vanderbilt.edu   \\
\\
Submitted to: \> \textit{Reliability Engineering and System Safety} \\
\>  May 2018\\

\bigskip
\end{tabbing}

\clearpage