### 2018-05-26 Paper/ack.tex
..

\section*{Acknowledgement}

M. Vohra and S. Mahadevan gratefully acknowledge funding support from the
National Science Foundation (Grant No. 1404823, CDSE Program). M. Vohra
would also like to sincerely thank Dr. Xun Huan at Sandia National Labs for
his guidance pertaining to the usage of TChem for the chemical kinetics application
in this work.