### 2018-05-28 Paper/disc.tex
..

\section{Summary and Conclusion}
\label{sec:disc}

%\begin{enumerate}
%\item Methodology is agnostic to the choice of surrogate.
%\item Outcome of parameter screening is QoI dependent. 
%\item As required for a PCE, the uncertain parameters need not be independent. 
% When working with a computational budget, consider relaxing the tolerance. 
% Higher the dimension reduction, greater the gains. However, in case of intensive models, even a small reduction could be
%advantageous. 
% In general, computational advantage is not guaranteed. Nevertheless, the proposed methodology is promising. 
% Gains with a reduced order surrogate are multiplied several time when evaluating the posterior distribution of the parameters
%since several surrogates are needed. 
%\end{enumerate}

In this work, we have presented a systematic approach for constructing a reduced-order
surrogate for scientific and engineering applications. Dimension reduction is accomplished
by identifying uncertain parameters that contribute relatively less towards the uncertainty
in the quantity of interest. These parameters deemed as \textit{unimportant} are determined
using a screening metric, defined in~\eqref{eq:cmu}. Initially, the metric is estimated
using model evaluations at a small set of samples in the parameter domain. However, these
estimates are refined by subsequent enrichment of the sample set during the screening
procedure presented in Algorithm~\ref{alg:screen}. The outcome of parameter screening is
assessed for the scope of dimension reduction. In a favourable scenario, a reduced-order
surrogate (ROS) is constructed. The ROS is tested for accuracy in a least-squares sense
as well as a probabilistic sense using a validation test suite. In the proposed framework,
a surrogate in the full-space (FSS) is constructed in tandem with parameter screening using
the available set of model evaluations. Both, ROS and FSS are
constructed using regression-based sparse basis methods. Simultaneous construction
of the FSS ensures that the computational effort associated with the proposed framework
does not overshoot the effort required to construct the FSS directly.
Hence, the ROS is constructed only in situations where computational gains are
expected.

Parameter screening methodology was implemented to low-to-moderate dimensional test problems
and an accurate ROS was constructed to demonstrate potential for computational gains in
each case. Furthermore, the overall framework was implemented  to a relatively higher 
dimensional application involving kinetics of the H$_2$/O$_2$ reaction mechanism. 
Significant dimension reduction was accomplished for two different scenarios involving
a fuel-rich and a fuel-lean mixture. In both cases, the resulting ROS was able to capture
the input-output relationship as well as the uncertainty in the quantity of interest with 
reasonable accuracy. Additional highlights of the proposed framework are as follows:
%\alennote{This is unclear: if we list limitations, we should be specific and
%not mix them with highlights;
%as for limitations, we can mention 
%e.g., we do not account for correlations, and our method applies only to 
%models with differentiable dependence on parameters, etc.; I would not make
%a bulleted list; I would briefly discuss the limitations, and if possible turn 
%them into strengths! e.g., we consider uncorrelated parameters, but that's okay as in 
%many cases correlations are not well understood and assuming independence is reasonable, etc.}
\begin{enumerate}
\item The proposed framework is agnostic to the choice of the surrogate although PCE's were used
in this work. 
\item The scope for dimension reduction and associated computational gains are
dependent upon the application as well as the quantity of interest. 
%\item The quantity of interest must be differentiable with respect to each uncertain parameter
%for the derivative-based sensitivity measures in~\eqref{eq:mu} to be defined. 
%to ensure accurate estimation of the derivative-based sensitivity measures in~\eqref{eq:mu}.
\item Substantial computational gains are expected in situations involving compute-intensive
simulations even if the scope for dimension reduction is small. 
%Hence, careful judgment 
%pertaining to the construction of a reduced-order surrogate as well as resource allocation is needed. is required when implementing the framework. 
\item Exponential gains can be realized in situations where multiple surrogates need to be
constructed as illustrated in the kinetics application. Other possible scenarios may
include inverse problems involving parameter estimation in a Bayesian setting. 
\end{enumerate}

Based on the results presented for a variety of test problems, 
and a higher-dimensional chemical kinetics application in this work, it can be said that the
proposed framework seems quite promising in its potential for identifying favourable
scenarios for constructing a reduced-order surrogate in a computationally efficient manner. 
However, it is important to remain cognizant about the underlying limitations of the
framework. For instance, it does not account for the existence of possible correlation 
between the uncertain parameters of the model, although in many cases, correlations
between parameters are not well understood, and assuming mutual independence 
would be reasonable. Similarly, the quantity of interest is required to be differentiable
with respect to each parameter in the considered domain. This condition, however, must
anyways be satisfied for the resulting surrogate to be able to represent the input-output
relationship for a given model with reasonable accuracy. Finally, careful assessment 
and decision-making is required with regards to proceeding with the
construction of a reduced-order surrogate based on the outcome of parameter
screening, and accuracy of the most-recent full-space surrogate as well as 
allocation of computational resources for parameter screening, and constructing the
validation test suite. Implementation to more advanced
applications in the future would enable further development of the proposed framework.