### 2018-05-28 Paper/disc.tex
minor ...

\section{Summary and Conclusion}
\label{sec:disc}

%\begin{enumerate}
%\item Methodology is agnostic to the choice of surrogate.
%\item Outcome of parameter screening is QoI dependent. 
%\item As required for a PCE, the uncertain parameters need not be independent. 
% When working with a computational budget, consider relaxing the tolerance. 
% Higher the dimension reduction, greater the gains. However, in case of intensive models, even a small reduction could be
%advantageous. 
% In general, computational advantage is not guaranteed. Nevertheless, the proposed methodology is promising. 
% Gains with a reduced order surrogate are multiplied several time when evaluating the posterior distribution of the parameters
%since several surrogates are needed. 
%\end{enumerate}

In this work, we have presented a systematic approach for constructing a reduced-order
surrogate for scientific and engineering applications. Dimension reduction is accomplished
by identifying uncertain parameters that contribute relatively less towards the uncertainty
in the quantity of interest. These parameters deemed as \textit{unimportant} are determined
using a screening metric, defined in~\eqref{eq:cmu}. Initially, the metric is estimated
using model evaluations at a small set of samples in the parameter domain. However, these
estimates are refined by subsequent enrichment of the sample set during the screening
procedure presented in Algorithm~\ref{alg:screen}. The outcome of parameter screening is
assessed for the scope of dimension reduction. In a favorable scenario, a reduced-order
surrogate (ROS) is constructed. The ROS is tested for accuracy in a least-squares sense
as well as a probabilistic sense using a validation test suite. In the proposed framework,
a surrogate in the full-space (FSS) is constructed in tandem with parameter screening using
the available set of model evaluations. Both, ROS and FSS are
constructed using regression-based sparse basis methods. Simultaneous construction
of the FSS ensures that the computational effort associated with the proposed framework
does not overshoot the effort required to construct the FSS directly.
Hence, the ROS is constructed only in situations where computational gains are
expected.

Parameter screening methodology was implemented to low-to-moderate dimensional
test problems and an accurate ROS was constructed to demonstrate potential for
computational gains in each case. Furthermore, the overall framework was
implemented  to a relatively higher dimensional application involving kinetics
of the H$_2$/O$_2$ reaction mechanism.  Significant dimension reduction was
accomplished for two different scenarios involving a fuel-rich and a fuel-lean
mixture. In both cases, the resulting ROS was able to capture the input-output
relationship as well as the uncertainty in the quantity of interest with
reasonable accuracy. Additional highlights of the proposed framework are as
follows:
\begin{enumerate}
\item The proposed framework is agnostic to the choice of the surrogate although PCE's were used
in this work. 
\item The scope for dimension reduction and associated computational gains are
dependent upon the application as well as the quantity of interest. 
%\item The quantity of interest must be differentiable with respect to each uncertain parameter
%for the derivative-based sensitivity measures in~\eqref{eq:mu} to be defined. 
%to ensure accurate estimation of the derivative-based sensitivity measures in~\eqref{eq:mu}.
\item Substantial computational gains are expected in situations involving
compute-intensive simulations even if the scope for dimension reduction is
small. 
%Hence, careful judgment pertaining to the construction of a reduced-order
%surrogate as well as resource allocation is needed. is required when
%implementing the framework.

\item Exponential gains can be realized in situations where multiple surrogates need to be
constructed as illustrated in the kinetics application. Other possible scenarios may
include inverse problems involving parameter estimation in a Bayesian setting. 
\end{enumerate}

Based on the results presented for a variety of test problems, and a
higher-dimensional chemical kinetics application in this work, it can be said
that the proposed framework seems quite promising in its potential for
identifying favorable scenarios for constructing a reduced-order surrogate in a
computationally efficient manner.  However, it is important to remain cognizant
about the limitations of the framework as well. For instance, it does not
account for the existence of possible correlations between the uncertain
parameters of the model. However, while the assumption of independent
parameters is not always justified, in many cases, correlations between
parameters are not well understood a priori, and assuming mutual independence
could be reasonable at least in an initial screening stage using DGSMs.
Additionally, the quantity of interest is required to be differentiable with
respect to each parameter in the considered domain. This condition, however,
must anyways be satisfied for the resulting surrogate to be able to represent
the input-output relationship for a given model with reasonable accuracy.
Finally, careful assessment and decision-making is required with regards to
proceeding with the construction of a reduced-order surrogate based on the
outcome of parameter screening, and accuracy of the most-recent full-space
surrogate as well as allocation of computational resources for parameter
screening, and constructing the validation test suite. Implementation to more
advanced applications in the future would enable further development of the
proposed framework.