### 2018-06-12 Paper/abstract.tex
..

\section*{Abstract}

In this effort, we develop a systematic approach to constructing a surrogate
in a reduced input space using derivative-based global sensitivity
measures~(DGSMs)~\cite{Sobol:2010}. A sparse set of model evaluations in
the original input space is used for an approximate evaluation of DGSMs
for individual uncertain inputs of the model. An iterative screening
procedure is developed to exploit DGSM estimates in order to identify
the `unimportant' inputs. The screening procedure forms an integral part
of an overall framework for adaptive construction of a surrogate in the
reduced-space. The framework is tested for computational efficiency through
an initial implementation in simple test cases such as the classic Borehole
function, and a semilinear elliptic PDE with a random source term. 
Motivated by these experiences, the framework is further implemented to
a 19-dimensional problem with uncertain rate constants associated with
the H$_2$/O$_2$ reaction mechanism~\cite{Yetter:1991}. It is observed
that significant computational gains can be accomplished by constructing
reliable and accurate low-dimensional surrogates using the proposed 
framework for practical problems.