### 2018-06-15 Paper/ack.tex
..

\section*{Acknowledgement}

M. Vohra and S. Mahadevan gratefully acknowledge funding support from the
National Science Foundation (Grant No. 1404823, CDSE Program). 
C. Safta was supported by the U.S. Department of Energy, Office of Science,
Basic Energy Sciences, as part of the Computational Chemical Sciences Program.
M. Vohra would also like to sincerely thank Dr. Xun Huan at Sandia National Labs for
his guidance pertaining to the usage of TChem for the chemical kinetics application
in this work. Sandia National Laboratories is a multimission laboratory managed 
and operated by National Technology \& Engineering Solutions of Sandia, LLC, 
a wholly owned subsidiary of Honeywell International Inc., for the 
U.S. Department of Energy's National Nuclear Security Administration under contract DE-NA0003525. 
The views expressed in the article do not necessarily represent the
views of the U.S. Department Of Energy or the United States Government.