### 2018-10-14 Paper/revision/reviewer_replies.tex
replies ...

\documentclass[11pt,final]{article}
\renewcommand*\familydefault{\sfdefault}
%\usepackage{amssymb,amsmath,amsfonts,comment}
%\usepackage{amsmath,amssymb,graphicx,subfigure,psfrag}
\usepackage{amsmath,amssymb,graphicx,subfigure,psfrag,upgreek}
\usepackage{algorithm,algorithmic}
\usepackage{amssymb,mathrsfs}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{color,pdfcolmk}
\newcommand{\todo}[1]{\noindent\emph{\textcolor{red}{Todo: #1\:}}}
%\newcommand{\alennote}[1]{\noindent\emph{\vspace{1ex}\textcolor{cyan}{Alen: #1\:}}\\[1ex]}
\newcommand{\referee}[1]{\vspace{.1ex}\noindent{\textcolor{blue}{#1}}}

\begin{document}

We thank the reviewers for their careful reading of our article 
and the helpful comments and suggestions.
Please find below point-by-point replies (in black) to your comments and
questions (which are reprinted in blue). To give you an overview of all the
changes in the paper, we also provide a diff-document that highlights the
changes between the initial submission and this re-submission.\\[1ex]

\section{Replies to reviewer \#1}
\referee{This draft proposed a systematic approach for surrogate model construction in
reduced input parameter spaces in order to reduce the number of model
evaluations. The key idea is to approximate the screening parameter iteratively
in order to identify the unimportant inputs. The screening procedure also
integrates with the adaptive construction of a surrogate in the reduced space
to improve the efficiency of the methods. Several numerical examples are used
to demonstrate the efficiency and accuracy of the proposed framework.}

Comments and Questions:
\begin{enumerate}

\item \referee{P16, Figure.2,  the estimated screening parameters eq (5) based on eq(1)
quickly converge with a small number of samples (5-10 samples). It seems that
if $\mu_i$ is based on eq(1), it will still require a reasonable amount of
samples, can authors comment on why this example converges with a small number
of samples? The same question applies to Figure 5 and Figure 7.}

\item 
\referee{P17, Figure 3, what's the degree of the PCE used in the example? The same
question applies to Figure 6 and Figure 8.}

\item
\referee{
P25, Figure 7, it seems that the rank of the parameters and the difference of
$\mu_s$ doesn't change too much over iterations. Does it indicate the initial
parameter screening  (line 1-6 in Algorithm 1) is good enough for ranking the
parameters? Can the authors plot the similar figure for other examples to see
if the iterative screening procedure does make a difference on the ranking the
parameters compared with the initial parameter screening step?}
\end{enumerate}

\section{Replies to reviewer \#2}

\referee{
An adaptive surrogate model construction approach in reduced input parameter
space is proposed in this paper. Instead of variance based sensitivity indices,
derivative based global sensitivity measures are employed to calibration of the
important model inputs, upon which surrogate models are constructed. To
demonstrate the computational advantage, the proposed approach is then
validated against three examples.  The reviewer’s recommendation is ``major
revision''.}

Major comments:
\begin{enumerate}

\item
\referee{
In this paper, the constructed surrogate model is either a RSS or a FSS. Based
on the diagram on page 12, the resulting RSS is built upon the not sufficiently
accurate FSS model (which could lead the accuracy to decrease further), or
is the RSS constructed directly from the original model with reduced input
parameters? The diagram is ambigous on this issue.}

\item
\referee{The reviewer would also like to see the comparison of the computational cost
(not just the numerical results) of DGSMs to variance based sensitivity indices
with increasing parameter dimension, e.g. Sobol’ indices incorporated with
sparse grid, since the difference decreases with increasing number of parameters.}

\item 
\referee{In section 4 and 5, no predictive convergence using the test data sample is
presented, only the convergence of $\epsilon_\text{LOO}$, which reuses the training data as
the validation set, is shown. This shows only the convergence of the model,
and does not necessarily indicate the (predictive) accuracy of the constructed
model, hence could lead to false conclusion. The reviewer would like to see the
convergence of $\epsilon_{L_2}$.}

\item 
\referee{How would the proposed approach behave when presented with severe nonlinearity
or rare events where the screening metrics could misrepresent the significance
of parameters?}

\item 
\referee{The reviewer would like to see some comments on the change of computational
cost of the proposed approach with respect to the increase of parameter dimensions.}
\end{enumerate}

Minor comments have been addressed in the revised draft. Thanks for pointing them out.

\end{document}