### 2018-10-24 Paper/app.tex
..

\section{Application: H$_2$/O$_2$ Reaction Kinetics}
\label{sec:app}

%Problem set-up: something about the reaction - why is it important?
%different reactions, reaction rate definition, uncertain parameters,
%quantity of interest. 
%
%Implementation of the proposed methodology for two scenarios: lean
%mix and rich mix. Describe what we mean by lean and rich using
%stoichiometry. 
%
%Define all the inputs as per the framework and illustrate its
%implementation to this application for both scenarios

The proposed framework in section~\ref{sec:method} is implemented to 
the H$_2$/O$_2$ reaction mechanism provided in~\cite{Yetter:1991}.
The H$_2$/O$_2$ reaction is gaining a lot of attention as a potential
source of clean energy for applications such as 
transportation~\cite{Das:1996} and fuel cell 
applications~\cite{Loges:2008,Cosnier:2016}. \rebut{We begin by providing
the necessary background information for setting up the problem
in~\ref{sub:setup}. Results and discussion based on the implementation
of the proposed framework are provided in~\ref{sub:imp}. Finally, 
in~\ref{sub:cost}, we
provide a comparative analysis of the computational cost involved
in determining relative parameter importance using DGSMs and the
total-effect Sobol index as the dimensionality increases from 19 to 33
for the present application.}

\subsection{Problem Set-up}
\label{sub:setup}  
The mechanism comprises of 19 reactions including chain reactions,
dissociation/recombination reactions, and formation and consumption
of intermediate species as provided below in Table~\ref{tab:kinetics}.

\begin{table}[htbp]
\renewcommand*{\arraystretch}{1.2}
\begin{center}
\begin{tabular}{llll}
\toprule
Reaction \#     & Reaction &&\\
\bottomrule
$\mathcal{R}_1$ & H + O$_2$          & $\rightleftharpoons$ & O + OH \\
$\mathcal{R}_2$ & O + H$_2$          & $\rightleftharpoons$ & H + OH \\
$\mathcal{R}_3$ & H$_2$ + OH         & $\rightleftharpoons$ & H$_2$O + H \\
$\mathcal{R}_4$ & OH + OH            & $\rightleftharpoons$ & O + H$_2$O \\
$\mathcal{R}_5$ & H$_2$ + M          & $\rightleftharpoons$ & H + H + M \\
$\mathcal{R}_6$ & O + O + M          & $\rightleftharpoons$ & O$_2$ + M \\
$\mathcal{R}_7$ & O + H + M          & $\rightleftharpoons$ & OH + M \\
$\mathcal{R}_8$ & H + OH +M          & $\rightleftharpoons$ & H$_2$O + M \\
$\mathcal{R}_9$ & H + O$_2$ + M      & $\rightleftharpoons$ & HO$_2$ + M \\
$\mathcal{R}_{10}$ & HO$_2$ + H      & $\rightleftharpoons$ & H$_2$ + O$_2$ \\
$\mathcal{R}_{11}$ & HO$_2$ + H      & $\rightleftharpoons$ & OH + OH \\
$\mathcal{R}_{12}$ & HO$_2$ + O      & $\rightleftharpoons$ & O$_2$ + OH \\
$\mathcal{R}_{13}$ & HO$_2$ + OH     & $\rightleftharpoons$ & H$_2$O + O$_2$ \\
$\mathcal{R}_{14}$ & HO$_2$ + HO$_2$ & $\rightleftharpoons$ & H$_2$O$_2$ + O$_2$ \\
$\mathcal{R}_{15}$ & H$_2$O$_2$ + M  & $\rightleftharpoons$ & OH + OH + M \\
$\mathcal{R}_{16}$ & H$_2$O$_2$ + H  & $\rightleftharpoons$ & H$_2$O + OH \\
$\mathcal{R}_{17}$ & H$_2$O$_2$ + H  & $\rightleftharpoons$ & HO$_2$ + H$_2$ \\
$\mathcal{R}_{18}$ & H$_2$O$_2$ + O  & $\rightleftharpoons$ & OH + HO$_2$ \\
$\mathcal{R}_{19}$ & H$_2$O$_2$ + OH & $\rightleftharpoons$ & HO$_2$ + H$_2$O \\
\bottomrule
\end{tabular}
\end{center}
\caption{Reaction mechanism for H$_2$/O$_2$ from~\cite{Yetter:1991}}.
\label{tab:kinetics}
\end{table}

The reaction rate for the $i^{th}$ reaction as a function of temperature
is given as follows:
\be
k_i(T) = A_iT^{n_i}\exp(-E_{a,i}/RT), 
\label{eq:rate}
\ee
%
where $A_i$ is the pre-exponent, $n_i$ is the index of $T$, $E_{a,i}$ is the
activation energy corresponding to the $i^{th}$ reaction, and $R$ is the
universal gas constant. 
The TChem~\cite{Safta:2011} software package is used to model homogeneous
ignition at constant pressure for a range of initial conditions for the
fuel-oxidizer mixture. During the simulation, the fuel-oxidizer mixture goes 
through a radical build-up phase followed by a sharp increase in temperature
as heat is released during the thermal runaway. We focus on quantifying the 
uncertainty in the \emph{ignition delay} due to uncertainty associated 
with the pre-exponent, $A_i$, for each reaction. The ignition delay 
is defined as the inflection point on the temperature profile during the thermal
runaway. The total number of uncertain parameters in the
present case is 19.  The $A_i$'s are considered to be uniformly distributed in
the interval: $[0.9A_i^\ast, 1.1A_i^\ast]$; $A_i^\ast$ being the nominal
estimate corresponding to the $i^{th}$ reaction.
The set of nominal values used in the computations, for parameters 
in~\eqref{eq:rate} are provided
in~\cite{Yetter:1991}.

While the dimensionality of the problem is relatively moderate,
constructing a surrogate in the 19-dimensional parameter space could still be
expensive. Hence, we explore the possibility of constructing a
reduced-space surrogate (RSS) using the framework presented 
in section~\ref{sec:method}. 
In the present study, we focus on two scenarios: fuel(H$_2$)-rich, and
fuel(H$_2$)-lean. Consider the global reaction:
%
\be
2\text{H}_2 + \text{O}_2 \rightarrow 2\text{H}_2\text{O}
\label{eq:global}
\ee 
%
The equivalence ratio $\phi$ is defined as follows:
%
\be
\phi = \frac{(M_{\text{H}_2}/M_{\text{O}_2})_\text{obs}}{(M_{\text{H}_2}/M_{\text{O}_2})_\text{st}}
\label{eq:phi}
\ee
%
The numerator in the right-hand-side represents the observed (obs) fuel-oxygen
mass ratio at a given condition and the denominator represents the
stoichiometric (st) ratio of the same quantity. Hence, $\phi$ = 1 at
stoichiometric conditions. The equivalence ratio can be altered by changing the
amount of O$_2$ in the mixture. In the case of a lean
mixture,~\eqref{eq:global} can be written as follows:
%
\be
2\text{H}_2 + \alpha\text{O}_2 \rightarrow 2\text{H}_2\text{O} + (\alpha-1)\text{O}_2 
\hspace{3mm} (\alpha>1)
\label{eq:lean}
\ee 
%
Similarly, for the case when the mixture if fuel rich,~\eqref{eq:global} is modified
as follows:
%
\be
2\text{H}_2 + \alpha\text{O}_2 \rightarrow 2\alpha\text{H}_2\text{O} + 2(1-\alpha)\text{H}_2
\hspace{3mm} (\alpha<1)
\label{eq:rich}
\ee 
%
Eqs.~\eqref{eq:lean} and~\eqref{eq:rich} can be generalized as follows:
%
\be
2\text{H}_2 + \alpha\text{O}_2 \rightarrow 2\min(1,\alpha)\text{H}_2\text{O} + 
\max(\alpha-1,0)\text{O}_2 + \max(0,2-2\alpha)\text{H}_2
\label{eq:gen}
\ee 
%
From the above set of chemical equations, the relationship between $\phi$
and $\alpha$ can be easily obtained as $\phi~=~\frac{1}{\alpha}$.
Since $\phi>1$ corresponds to a rich mixture, and $\phi<1$ corresponds to a
lean mixture, we consider $\phi$ = 2.0 and 0.5 to investigate the two scenarios
respectively.

\subsection{DGSM-guided surrogate construction}
\label{sub:imp}
We apply the parameter screening algorithm with the following
parameters: $\tau_\text{screen}$, $s_\text{min}$,
$s_\text{max}$, $\beta$ are fixed at 0.2, 3, 10, and 1.0 respectively for both cases.
Additionally, the value of $\tau$ is considered to be 1.0$\times 10^{-17}$ and
5.0$\times 10^{-17}$ in the rich and lean case respectively. Such a small value
of $\tau$ for this application is a consequence of the nature of convergence exhibited
by the sensitivity measures. Moreover, the screening procedure is carried out
for atleast $s_\text{min}$ number of iterations in order to bolster our confidence
in the estimates.

Following the steps outlined in the flow-diagram in Figure~\ref{fig:flow}, model
evaluations are initially generated at $n_1$ = 5 samples. The evaluations are
used to construct a regression-based surrogate in the full-space. As
expected, the surrogate is found to be highly inaccurate. Moreover, unlike the
test problems in section~\ref{sec:examples}, we do not estimate the Sobol
total-effect sensitivity indices in the interest of following the overall framework
closely. Hence, we proceed to the screening step to estimate the screening metric
for the uncertain pre-exponents, $A_i$'s. Results are plotted below
in Figure~\ref{fig:sense_kinetics}~(top row) for both cases. Furthermore, we illustrate
the decay in the value of $\Delta\mu_s$ with iterations in 
Figure~\ref{fig:sense_kinetics}~(bottom row).

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_lean}
  \\ \vspace{2mm}
  \includegraphics[width=0.45\textwidth]{./Figures/mu_rich}
  \includegraphics[width=0.45\textwidth]{./Figures/mu_lean}
\caption{Top: Estimates of $\widehat{\mathcal{C}_i\mu_i}$ for $A_i$'s in the case
of fuel-rich mixture~(left) and fuel-lean mixture~(right). Bottom: The value of
$\Delta\mu_s$ during three iterations within the screening step are plotted for
the case of fuel-rich mixture~(left) and fuel-lean mixture~(right).}
\label{fig:sense_kinetics}
\end{center}
\end{figure}
%
The screening metric estimates in the above plots are observed to converge with
only a few samples (5--10). Moreover, out of the 19 uncertain pre-exponents,
only $A_1$, $A_9$, $A_{15}$, and $A_{17}$ seem to be important in the fuel-rich
case, whereas, only $A_1$, $A_9$, and $A_{15}$ seem important in the fuel-lean
case, based on the value of $\tau_\text{screen}$. These observations are indicative
of the potential for significant reduction in the dimensionality of this problem. 
A reduced-space
surrogate constructed using the proposed framework could thus lead to large computational
gains. The decay of $\Delta\mu_s$ with iterations is expected and builds our confidence
in the screening procedure in both cases.

A reduced-space surrogate (RSS) was constructed in 4D for the fuel-rich case,
and in 3D for
the fuel-lean case. Figure~\ref{fig:err_samples_kinetics}~(left) illustrates a
comparison of convergence characteristics for the PCEs constructed in the
full-space and the reduced-space for the fuel-rich case. Note that the
plot is generated using the implementation of least angle regression (LAR)
for sparse PCEs in UQLab.  
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/err_samples_kinetics}
   \includegraphics[width=0.44\textwidth]{./Figures/errL2_samples_kinetics}
\caption{Left: A semi-log plot of $\epsilon_\text{\tiny LOO}$ as a function of
number of model evaluations in the full-space (19D) and the reduced-space (4D)
for the fuel (H$_2$)-rich case i.e. $\phi$ = 2.0. \rebut{The degree of
the PCE constructed using 60 training points was found to be 1 and 3 in the 19D and the 4D cases respectively.
Right: Semilog plot of the relative L-2 error norm ($\epsilon_{\mbox{\tiny{L-2}}}$)
for the PCEs constructed in 4D and 19D.}}
\label{fig:err_samples_kinetics}
\end{center}
\end{figure}
%
The leave-one-out cross validation error is observed to drop initially
and plateau with the increase in training points for the 19-dimensional
PCE. However, in the case of 4-dimensional PCE, the error exhibits a
monotonic behavior and is found to be smaller than $\mathcal{O}(10^{-5})$
at 60 training points. Clearly, the RSS shows a much faster rate of convergence.
\rebut{Additionally, the
convergence of the relative error norm, $\epsilon_{\mbox{\tiny{L-2}}}$ in 
Figure~\ref{fig:err_samples_kinetics}~(right) indicates that for the case of 10
samples, the 4D PCE is observed to be more accurate. However, the 19D PCE
is observed to be more accurate by an order of magnitude as the sample size
increases. Specifically, the 4D PCE was found to be accurate within 1.8$\%$ in the
fuel-rich case, and within 3.1$\%$ in the fuel-lean case. Therefore, the 4D PCE
is still observed to be reasonably accurate in this case. Moreover, its faster convergence
as mentioned earlier should help reduce computational effort associated with
surrogate model construction.}
Similar trends (not included) were observed in the fuel-lean case.

Model evaluations at 1000 samples in the test suite
are further used to plot a normalized histogram of the ignition time in 
Figure~\ref{fig:pdf_kinetics}. To verify the accuracy of the RSS in a 
probabilistic-sense, we compare the histogram plot with a PDF of ignition time
using surrogate predictions at 10$^{6}$ samples in the reduced space in both cases. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/pdf_comp_rich}
  \includegraphics[width=0.45\textwidth]{./Figures/pdf_comp_lean}
\caption{A normalized histogram based on model evaluations at 1000 samples is plotted
along with a PDF of ignition delay for the fuel-rich case (left) and the fuel-lean
case (right).}
\label{fig:pdf_kinetics}
\end{center}
\end{figure}
%
Clearly, the RSS captures the spread as well as the modal estimate of
the ignition delay in both scenarios. Hence, the proposed framework 
has enabled significant dimension reduction and construction of an accurate
RSS for multiple scenarios pertaining to the H$_2$/O$_2$ reaction
mechanism.

\rebut{
\subsection{Comparative cost analysis}
\label{sub:cost}

In this section, we perform a comparative analysis of the computational cost
associated with obtaining converged estimates of parametric sensitivities
using DGSM, and the total-effect Sobol indices ($\mathcal{T}(\theta_i)$). The former is computed
using the parameter screening algorithm (Algorithm~\ref{alg:screen}), and the latter is
estimated analytically using a sparse-basis PCE. Note that estimating $\mathcal{T}(\theta_i)$
using a PCE has been shown to be more efficient compared to sampling techniques.
Moreover, a sparse-basis PCE is expected to require fewer training points which
enhances computational savings. To investigate the impact of the dimensionality, the
analysis is performed with 19 uncertain parameters, i.e.
the pre-exponents~($A_i$'s) as discussed earlier in~\ref{sub:setup} and~\ref{sub:imp}, and
33 uncertain parameters wherein the activation energies~($E_{a,i}$'s) are also considered to
be uncertain. The $E_{a,i}$'s are also considered to be uniformly distributed in the
interval: [$0.9E_{a,i}^\ast,1.1E_{a,i}^\ast$], where $E_{a,i}^\ast$ is the nominal value
for the $i^\text{th}$ reaction as provided in~\cite{Yetter:1991}. Note that only those
$E_{a,i}$'s with a non-zero nominal value are considered as uncertain, and therfore the
total number of uncertain parameters in the high-dimensional case is 33 as opposed to 38.

The 33-dimensional sparse basis PCE was assessed for its convergence characteristics
and its predictive accuracy using $\epsilon_{\mbox{\tiny LOO}}$~(see~\eqref{eq:loo}) and
$\epsilon_{\mbox{\tiny{L-2}}}$~(see~\eqref{eq:l2}) respectively as shown in
Figure~\ref{fig:err_samples}.  
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/err_samples_kinetics33D}
\caption{A semi-log plot of $\epsilon_\text{\tiny LOO}$  and $\epsilon_{\mbox{\tiny{L-2}}}$
as a function of the number of training points in the 33D parameter space 
for the fuel (H$_2$)-rich case i.e. $\phi$ = 2.0. The degree of
the PCE constructed using 500 training points was found to be 3.}
\label{fig:err_samples}
\end{center}
\end{figure}
%
The sparse-basis PCEs constructed in the 19 and the 33 dimensional parameter
space are considered to have converged once $\epsilon_\text{\tiny LOO}$ is found to be
smaller than 6.0$\times$10$^{-3}$. As observed in Figures~\ref{fig:err_samples_kinetics}
and~\ref{fig:err_samples}, the number of training points required in the 19D
and the 33D cases are found to be 60 and 500 respectively. As expected, the predictive
accuracy is observed to increase with increase in the number of training points in both cases.
The estimate of $\epsilon_{\mbox{\tiny{L-2}}}$ was found to be 7.22$\times$10$^{-2}$ using the
converged 33D PCE and an independent set of model evaluations at 1000 samples.

Sensitivity estimates for the 33 uncertain parameters are hence determined
using the DGSM-based strategy and by evaluating $\mathcal{T}(\theta_i)$ analytically
using the sparse-basis PCE as discussed earlier in this section. The estimates based
on the two approaches were plotted separately and placed side-by-side for comparison
in Figure~\ref{fig:sense33D}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.42\textwidth]{./Figures/ub33D}
  \includegraphics[width=0.40\textwidth]{./Figures/sens_kinetics33D}
\caption{
Left: Estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$) for each 
uncertain parameter, obtained using $N$ = 5, 10, 15, and 20 samples in the full parameter space.
Right: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain parameters in the
semilinear elliptic PDE~\eqref{eq:semilinear}.}
\label{fig:sense33D}
\end{center}
\end{figure}
%

}